---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  eval=FALSE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# afdscraper

<!-- badges: start -->
<!-- badges: end -->

afdscraper is an r package for scraping data from the [Australian Faunal Directory](https://biodiversity.org.au/afd/home)

## Installation

You can install the development version of afdscraper from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("alexpiper/afdscraper")
```

# Working examples

afdscraper can be used for a number of different applications

## Retrieving checklist data for large taxonomic groups

One of the main uses of afdscraper is to retrieve CSV data for a large amount of taxa, overcoming the current 1000 taxon limit on the website.

```{r fetch CSV}
# Load the afdscraper package
library(afdscraper)

# Also load tidyverse packages for their handy filtering and summarising functiosn
library(tidyverse)

# Download all checklist data for Insecta
afd_insecta <- fetch_afd_checklist("Insecta", retry_attempt=3, retry_wait=5, quiet=FALSE)

head(afd_insecta)
```


## Checking presence of a taxon in Australia

The presence of a species on AFD can be checked. This is useful for automated cross referencing of species detections to identify potential new species introductions.

```{r check presence}
check_afd_presence("Bactrocera tryoni")
```

## Get the taxonomic rank of a query taxon

The taxonomic rank of a query taxon name can be obtained. 

```{r tax rank}
get_afd_rank("Scaptodrosophila")
```

## Retrieve the immediate children of a query taxon

The taxonomic children can be retrieved for taxonomic ranks above species.

```{r taxonomic children}
get_afd_children("Tephritidae")
```

## Retrieve distribution information

Distribution data can be retrieved for taxa, either by states (type = "states"), or by [Australia's Bioregions](https://www.environment.gov.au/land/nrs/science/ibra) (type = "ibra"). Note that occurance data is lacking for many of the taxa listed on AFD.

```{r dist info}
get_afd_occurance("Bactrocera tryoni", type = "states") 
```

### Multiple taxa

All of the above functions can also handle multiple taxa provided to the function or read in from a file.

```{r multiple taxa}
taxa <- c("Bactrocera tryoni", "Scaptodrosophila", "Nitidulidae")
check_afd_presence(taxa)

# Read in a list from a file
taxa <- readLines("taxon_list.txt")
check_afd_presence(taxa)
```
